% $LAAS: getting.tex 2010/06/23 15:03:35 mallet $
%
% Copyright (c) 2009-2010 LAAS/CNRS
% All rights reserved.
%
% Permission to use, copy, modify, and distribute this software for any purpose
% with or without   fee is hereby granted, provided   that the above  copyright
% notice and this permission notice appear in all copies.
%
% THE SOFTWARE IS PROVIDED "AS IS" AND THE AUTHOR DISCLAIMS ALL WARRANTIES WITH
% REGARD TO THIS  SOFTWARE INCLUDING ALL  IMPLIED WARRANTIES OF MERCHANTABILITY
% AND FITNESS. IN NO EVENT SHALL THE AUTHOR  BE LIABLE FOR ANY SPECIAL, DIRECT,
% INDIRECT, OR CONSEQUENTIAL DAMAGES OR  ANY DAMAGES WHATSOEVER RESULTING  FROM
% LOSS OF USE, DATA OR PROFITS, WHETHER IN AN ACTION OF CONTRACT, NEGLIGENCE OR
% OTHER TORTIOUS ACTION,   ARISING OUT OF OR IN    CONNECTION WITH THE USE   OR
% PERFORMANCE OF THIS SOFTWARE.
%
%                                             Anthony Mallet on Sat Jan 10 2009
%


\section{Where to get robotpkg and how to keep it up-to-date} % ------------
\label{section:getting}

Before you download and extract the files, you need to decide where you want to
extract them and  where you want robotpkg to install  packages.  By defaut, the
{\tt  /opt/openrobots} directory is  used. In  the rest  of this  document, the
installation path is called the {\em prefix}.

{\tt robotpkg}  will {\em never}  require administration privileges  by itself.
We thus recommend that you do not install or run robotpkg as the root user.  If
something ever goes really  wrong, it might go less wrong if  it is not running
as root. If you want to  install to the default location {\tt /opt/openrobots},
we recommend that you create this directory owned by a regular user.

Creating or using {\tt /opt/openrobots} typically requires administration ({\em
a.k.a.}  ``{\tt root}'') privileges.  If  you don't have such privileges (or if
you want  to install to a different  location), you have to  unpack the sources
and  install the  binary packages  in  another prefix.  If you  don't have  any
special administration  rights on the target  machine, a safe bet  is to choose
the {\tt  \${HOME}/openrobots} location, as  the {\tt \${HOME}}  directory will
always be writable by yourself.

Any prefix  will work, but please  note that you should  choose an installation
path which is dedicated to robotpkg packages and not shared with other programs
(e.g., we do  not recommend to use  a prefix of {\tt /usr}).   Also, you should
not try to add any of your  own files or directories (such as {\tt src/}) below
the prefix  tree.  This  will prevent possible  conflicts between  programs and
other files  installed by the  package system and  whatever else may  have been
installed there.

Finally,  the  installation  path   shall  not  contain  white-space  or  other
characters that are interpreted specially by the shell and some other programs:
use only letters, digits, underscores and dashes.

The rest of this document will  assume that you are using {\tt /opt/openrobots}
as the prefix. You should adapt this path to whatever prefix you choosed.


\subsection{Getting the binary bootstrap kit}

At the moment, the  binary bootstrap kit is not  available. Please get the {\tt
robotpkg} sources as described in the next section.


\subsection{Getting robotpkg for source compilation}

{\tt         robotpkg}    sources   are       distributed     {\em    via}  the
\href{http://git-scm.com/}{\tt git}  software  content management system.  {\tt
git} will probably be readily available on your system but if you don't have it
installed   or if you    are  unsure  about it,   contact  your  local   system
administrator.

There are two download methods: the anonymous one and the authenticated
one:

\begin{itemize}

  \item  Anonymous download is the  recommended method if  you  don't intend to
  work on  the   robotpkg infrastructure  itself,  nor commit   any  changes or
  packages  additions  back to  the  robotpkg main repository. Furthermore, the
  possibility to send contributions via patches is still open.

  As your regular user, simply run in a shell:

\begin{verbatim}
% cd /opt/openrobots
% git clone http://softs.laas.fr/git/robots/robotpkg.git
\end{verbatim}


  \item Authenticated   download requires a  valid  login on the  main robotpkg
  repository, and will give you full commit access  to this repository.  Simply
  run the following:

\begin{verbatim}
% cd /opt/openrobots
% git clone ssh://softs.laas.fr/git/robots/robotpkg
\end{verbatim}

\end{itemize}


\subsection{Keeping robotpkg up-to-date} % ---------------------------------

{\tt robotpkg} is   a   living thing:    updates  to  the packages are     made
perdiodicaly,  problems are fixed,   enhancements  are developed\ldots If   you
downloaded the robotpkg sources via git, you  should keep it up-to-date so that
you get the most  recent packages descriptions. This is done by running {\tt
git pull} in the robotpkg source directory:

\begin{verbatim}
% cd /opt/openrobots/robotpkg
% git pull
\end{verbatim}

When you update robotpkg, the git program will only  touch those files that are
registered in the git repository. That means that any packages that you created
on your own will stay unmodified. If you change  files that are managed by git,
later updates will try to merge your changes with those that  have been done by
others. See the {\tt git-pull} manual for details.

If you want  to be informed  of package additions  and other  updates, a public
mailing    list  is   available    for   your    reading   pleasure.  Go     to
\url{https://sympa.laas.fr/sympa/info/robotpkg}    for   more  information  and
subscription.
