% $LAAS: internal.tex 2009/01/16 23:58:51 tho $
%
% Copyright (c) 2009 LAAS/CNRS
% All rights reserved.
%
% Permission to use, copy, modify, and distribute this software for any purpose
% with or without   fee is hereby granted, provided   that the above  copyright
% notice and this permission notice appear in all copies.
%
% THE SOFTWARE IS PROVIDED "AS IS" AND THE AUTHOR DISCLAIMS ALL WARRANTIES WITH
% REGARD TO THIS  SOFTWARE INCLUDING ALL  IMPLIED WARRANTIES OF MERCHANTABILITY
% AND FITNESS. IN NO EVENT SHALL THE AUTHOR  BE LIABLE FOR ANY SPECIAL, DIRECT,
% INDIRECT, OR CONSEQUENTIAL DAMAGES OR  ANY DAMAGES WHATSOEVER RESULTING  FROM
% LOSS OF USE, DATA OR PROFITS, WHETHER IN AN ACTION OF CONTRACT, NEGLIGENCE OR
% OTHER TORTIOUS ACTION,   ARISING OUT OF OR IN    CONNECTION WITH THE USE   OR
% PERFORMANCE OF THIS SOFTWARE.
%
%                                             Florent Lamiraux on Sat Jan 24 2009
%

\subsection*{Using autotools}

The following variables and inclusion are useful for packages managed by {\tt autotools}. They should be defined in the installation Makefile of the package.

\paragraph{${\tt GNU\_CONFIGURE}$} should be defined in order to invoke configure script for configuration. 

\subsection*{Using cmake}

The following variables and inclusion are useful for packages managed by {\tt cmake}. They should be defined in the installation Makefile of the package.

\paragraph{${\tt USE\_CMAKE}$} should be set to {\tt yes} to invoke cmake for configuration. Moreover, the following inclusion should added after the definition of the above variable in the Makefile (assuming the Makefile is two directories below {\tt robotpkg} root directory):\\
{\tt include ../../mk/sysdep/cmake.mk}

\subsection*{Compiling fortran files}

If some packages include fortran files, it is important that all these files are compiled with the same compiler. To achieve so, it the following variable should be set in all packages including fortran files.

\paragraph{${\tt USE\_LANGUAGES+=fortran}$} consistently sets ${\tt FC}$ environment variable before configuring packages including fortran files. Thus, all fortran files will be compiled with the same compiler.